---
title: "Myxo notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(mclust)
library(tidyverse)
library(brms)
library(tidybayes)
library(cowplot)
library(ggforce)
library(mgcv)
library(modelr)
library(GGally)
theme_set(theme_cowplot())

library(furrr)
plan(multiprocess(workers = 3)) # More cores is faster, but some of these future_maps require a ton of memory per core.

current_script_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_script_dir)
```

Load data from CSV files, clean, and save as an RDS file. If rerun, check clusters.
```{r}
load_file <- function(path) {
  read_csv(path, col_types = cols()) %>%
    mutate(file_name = str_extract(string = path, pattern = "[a-zA-z0-9_]*\\.txt"))
}

load_dir <- function(dir) {
  file_list <- list.files(path = dir, pattern = "*.txt", full.names = T)
  map_df(file_list, load_file)
}

myxo_data <- load_dir("./csv/raw_data") %>%
  mutate(time_point = str_remove_all(str_extract(file_name, "_\\d\\dh_"),'_'), body_id = str_extract(file_name, "body_\\d+")) %>%
  filter(sphericity <= 1) %>%
  mutate(z_dist = WeightedCentroid_3 * 0.2,
         z_dist = ifelse(body_id == "body_3" & time_point == "30h", z_dist - 0.5, z_dist)) %>% # body_3 @ 30h is shifted relative to the others.
  filter(nb_branch_pt == 0, z_dist > 1,  z_dist < 9.5) %>%
  mutate(z_dist = ifelse(!(body_id %in% c("body_1", "body_2", "body_3")) & (z_dist > 4.5), NA, z_dist)) %>%
  filter(!is.na(z_dist))

# Reclustering. Don't re-run this unless necessary, or you'll have to re-determine the classification labels.
#
# myxo_data_subset <- myxo_data %>%
#   select(EquivDiameter, SurfaceArea, SkeletonSize, MeanWidth, time_point) %>%
#   mutate(logEquivDiameter = log(EquivDiameter),
#          sqrtSurfaceArea = sqrt(SurfaceArea),
#          logSqrtSurfaceArea = log(sqrtSurfaceArea),
#          logSkeletonSize = log(SkeletonSize),
#          logMeanWidth = log(MeanWidth)) %>%
#   select(-EquivDiameter, -SurfaceArea, -SkeletonSize, -MeanWidth, -sqrtSurfaceArea)
# myxo_clust <- myxo_data_subset %>%
#   select(-time_point) %>%
#   densityMclust(G = 4)
# write_rds(myxo_clust, './rds/mclust_model.rds', compress = 'gz')
#
# Determine the classification
# ggplot(myxo_data) +
#    geom_point(aes(x = EquivDiameter %>% log, y = sphericity, col = classification), alpha = 0.75) +
#   facet_grid(. ~ time_point)

myxo_data_subset <- myxo_data %>%
  select(EquivDiameter, SurfaceArea, SkeletonSize, MeanWidth, time_point) %>%
  mutate(logEquivDiameter = log(EquivDiameter),
         sqrtSurfaceArea = sqrt(SurfaceArea),
         logSqrtSurfaceArea = log(sqrtSurfaceArea),
         logSkeletonSize = log(SkeletonSize),
         logMeanWidth = log(MeanWidth)) %>%
  select(-EquivDiameter, -SurfaceArea, -SkeletonSize, -MeanWidth, -sqrtSurfaceArea, -time_point)

mapping <- list("1" = "debris", "2" = "spore", "3" = "transitioning", "4" = "rod")
clust_model <- read_rds('./rds/mclust_model.rds')
pred <- predict.Mclust(clust_model, myxo_data_subset)

myxo_data <- myxo_data %>%
  ungroup() %>%
  mutate(classification = pred$classification %>% as.factor(),
         labels = mapping[classification %>% as.character()] %>% unlist(),
         body_id = str_replace(body_id, "_(?=\\d$)", "_0"))

myxo_data <- myxo_data %>%
  group_by(body_id, time_point) %>%
  filter(labels != "debris") %>%
  mutate(labels_factor = factor(labels, levels = c("rod", "transitioning", "spore")),
         label_time = str_c(labels, time_point),
         dist_from_center = sqrt((WeightedCentroid_1 - mean(WeightedCentroid_1))^2 + (WeightedCentroid_2 - mean(WeightedCentroid_2))^2) * 0.1240,
         z_dist = WeightedCentroid_3 * 0.2)

write_rds(myxo_data, './rds/myxo_data.rds', compress = 'gz')
write_csv(myxo_data, './csv/myxo_data.csv')
```

Tables of cell counts per fruiting body, time point, etc.
```{r}
myxo_data <- read_rds("./rds/myxo_data.rds")

(counts <- myxo_data %>%
    ungroup() %>%
    group_by(body_id) %>%
    summarise(num_cells = n()))

write_csv(counts, "./csv/cell_counts_by_body_id.csv")

(counts_body_tp <- myxo_data %>%
    ungroup() %>%
    group_by(body_id, time_point) %>%
    summarise(num_cells = n()))

write_csv(counts_body_tp, "./csv/cell_counts_by_body_id_by_time_point.csv")

(counts_cell_type <- myxo_data %>%
    ungroup() %>%
    group_by(labels_factor) %>% 
    summarise(num_cells = n()))

write_csv(counts_cell_type, "./csv/cell_counts_by_body_cell_type.csv")
```


Figure 2D -- Cell type proportions over time
```{r}
cell_data_proportions <- read_rds("./rds/myxo_data.rds") %>%
  ungroup() %>%
  group_by(body_id, time_point, labels_factor) %>%
  summarize(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>%
  pivot_wider(names_from = labels_factor, values_from = n) %>%
  mutate(time_point = ordered(time_point, levels = c("27h", "30h", "36h", "48h")))

cell_data_proportions$count_vec <- with(cell_data_proportions, cbind(rod, transitioning, spore))

prop_model_path <- "./rds/model_rds/body_prop_model_all_bodies.rds"
prop_fn <- bf(count_vec | trials(total) ~ mo(time_point) + (mo(time_point) | body_id),
              muspore ~ mo(time_point) + (mo(time_point) | body_id),
              mutransitioning ~ mo(time_point) + (mo(time_point) | body_id),
              family = multinomial())

prop_model <- brm(prop_fn, data = cell_data_proportions, chains = 4, cores = 4,
                    warmup = 500, iter = 1500,
                    control = list("adapt_delta" = 0.99, "metric" = "diag_e", "max_treedepth" = 15))

write_rds(prop_model, prop_model_path, compress = "gz")

prop_model <- read_rds(prop_model_path)

n_total <- 1000

pred_by_body <- data_grid(cell_data_proportions, time_point, total = n_total) %>%
  add_fitted_draws(model = prop_model, re_formula = NA, summary = T, n = 4000) %>%
  mutate(time_point_numeric = time_point %>% str_extract("\\d\\d") %>% as.numeric,
         .value = .value/n_total)

(prop_model_all_bodies <- ggplot(pred_by_body) +
    stat_interval(aes(x = time_point_numeric, y = .value, group = .category, col = .category), point_interval = median_hdci, .width = c(0.90), alpha = 0.25, size = 7) +
    stat_summary(aes(x = time_point_numeric, y = .value, group = .category, col = .category), fun = median, geom = "point", size = 3) +
    labs(x = "Hours PS", y = "Proportion", col = "", fill = "") +
    theme(text = element_text(size = 18), axis.text = element_text(size = 18),
          strip.text.x = element_text(size = 18), legend.text = element_text(size = 18), legend.position = "top"))

ggsave("./figure_2/figure_2D_proportions_model_all_bodies.png", plot = prop_model_all_bodies, width = 6, height = 4, dpi = 300, bg = "white")
```



Figure 3A -- Radial distribution of cell classifications
Run radial density models
```{r}
cells_dist <- read_rds("./rds/myxo_data.rds") %>%
  filter(dist_from_center < 60)

const_area_annuli <- function(lb, ub, target_area) {
  r <- 0
  slice_vec <- c() 
  while (r < ub) {
    slice_vec <- c(slice_vec, r)
    r <- sqrt((target_area / pi) + r^2)
  }
  slice_vec
}

fitting_data <- tibble(r = cells_dist$dist_from_center,
                       label = cells_dist$labels,
                       body_id = cells_dist$body_id,
                       time_point = cells_dist$time_point,
                       dummy = 1)

gam_density <- function(filtered_data, label_str) {
  print(str_glue("Fitting density GAM to: {label_str}"))

  cell_counts_by_body <- filtered_data %>%
    group_by(body_id) %>%
    summarise(n = n())
  
  fd <- filtered_data %>%
    mutate(bin = cut(r, const_area_annuli(0, 65, (pi*65^2) / 20))) %>%
    group_by(body_id, bin) %>%
    summarise(n = n()) %>%
    mutate(left_pt = str_extract(bin %>% as.character, "\\d+(\\.\\d+)?(?=\\,)") %>% as.numeric,
           tot = map_dbl(body_id, ~filter(cell_counts_by_body, body_id == .x)$n),
           prop = n / tot,
           dummy = 1)
  
  (raw_points <- ggplot(fd) +
      geom_point(aes(x = left_pt, y = prop, col = body_id)) +
      geom_smooth(aes(x = left_pt, y = prop)))
  ggsave(filename = str_glue("./val/radial_density_smooth_{label_str}.png"), raw_points, width = 8, height = 6)
  
  formula <- bf(prop ~ t2(left_pt) + t2(left_pt, body_id, dummy, bs = 're'), family = Beta)
  
  fit <- brm(formula, data = fd, chains = 8, cores = 8, inits = 0, warmup = 250,
             iter = 5250, control = list("adapt_delta" = 0.999, max_treedepth = 14))
  
  write_rds(fit, path = str_glue("./rds/model_rds/radius_density/{label_str}.rds"), compress = "gz")
}

walk(list("27h", "30h", "36h", "48h"), ~filter(fitting_data, time_point == .x) %>% gam_density(str_glue("gam_density_{.x}")))
```

Figure 3A -- Radial distribution of cell classifications
Merge radial density models
```{r}
merge_radial_density <- function(filepath, name, run_all_body_ids = FALSE) {
  print(str_glue("Loading radial density model {filepath}"))
  fit <- read_rds(filepath)
  
  if (!run_all_body_ids) {
    body_ids = "body_01"
    dummy_val = 0
  } else {
    body_ids = unique(fit$data$body_id)
    dummy_val = 1
  }
  
  map_df(body_ids, function(bid) {
    pred_body <- data_grid(fit$data, left_pt = seq(0, 60, 0.2), body_id = bid, dummy = dummy_val) %>%
      add_fitted_draws(fit, n = 1000) %>%
      mutate(time_point = str_extract(name, "\\d\\dh"))
    return(pred_body)
  })
}

filename_regex <- "(?<=/)[A-Za-z_0-9\\.]*(?=\\.rds$)"

nobody_density_models <- map_df(list.files("./rds/model_rds/radius_density", pattern = ".*\\.rds", full.names = T),
                                ~ merge_radial_density(.x, str_extract(.x, filename_regex), FALSE))
write_rds(nobody_density_models, "./rds/model_rds/merged_density_models_radius_all_bodies.rds", compress = "gz")

all_body_density_models <- map_df(list.files("./rds/model_rds/radius_density", pattern = ".*\\.rds", full.names = T),
                                  ~ merge_radial_density(.x, str_extract(.x, filename_regex), TRUE))
write_rds(all_body_density_models, "./rds/model_rds/merged_density_models_radius_individual_bodies.rds", compress = "gz")
```

Figure 3A -- Radial distribution of cell classifications
Run radial proportions model
```{r}
cells_dist <- read_rds("./rds/myxo_data.rds")

fitting_data <- tibble(r = cells_dist$dist_from_center,
                       label = cells_dist$labels, 
                       body_id = cells_dist$body_id,
                       time_point = cells_dist$time_point,
                       dummy = 1)

gam_label <- function(filtered_data, label_str) {
  print(str_glue("Fitting label GAM to: {label_str}"))
  data <- filtered_data
  data$i <- with(data, cbind(label=='rod', label=='spore', label=='transitioning')) * 1
  colnames(data$i) <- c('rod','spore','transitioning')
  
  formula <- bf(i | trials(1) ~ 1,
                muspore ~ t2(r) + t2(r, body_id, dummy, bs='re'), 
                mutransitioning ~ t2(r) + t2(r, body_id, dummy, bs='re'),
                family = multinomial())
  
  fit <- brm(formula, data = data, chains = 8, cores = 8, inits = 0, warmup = 250,
             iter = 5250, control = list("adapt_delta" = 0.999, max_treedepth = 14))
  
  write_rds(fit, path = str_glue("./rds/model_rds/radius_label/{label_str}.rds"), compress = "gz")
}

walk(list("27h", "30h", "36h", "48h"), ~filter(fitting_data, time_point == .x) %>% gam_label(str_glue("gam_label_{.x}")))
```

Figure 3A -- Radial distribution of cell classifications
Merge radial proportions of all bodies (no body effect)
```{r}
merge_radial_label <- function(filepath, name) {
  print(str_glue("Loading radial label model {filepath}"))
  fit <- read_rds(filepath)
  
  model_out <- data_grid(data = fit$data, r = seq(0, 60, 0.5), body_id = "body_01", dummy = 0) %>%
    add_fitted_draws(fit, re_formula = NA, n = 4000) %>%
    mutate(body_id = body_id,
           name = name,
           time_point = str_extract(name, "\\d{2}h"))
  
  return(model_out)
}

all_models <- map_df(list.files("./rds/model_rds/radius_label", pattern = ".*\\.rds", full.names = T),
                     ~ merge_radial_label(.x, str_extract(.x, "(?<=/)[A-Za-z_0-9\\.]*(?=\\.rds$)")))

write_rds(all_models, "./rds/model_rds/merged_radial_proportions_all_bodies.rds", compress = "gz")
```

Figure 3A -- Radial distribution of cell classifications
Contrasts of radial distributions for hypotheses
```{r}
compare_radial_prop <- function(filepath, name, in_body_id) {
  print(str_glue("Loading {filepath}"))
  fit <- read_rds(filepath)
  
  # find point of maximal prob
  props <- data_grid(fit$data, r = seq(1, 60, 1), body_id = in_body_id, dummy = 0) %>%
    add_fitted_draws(fit, re_formula = NA, n = 4000)
  
  (validation_plot <- ggplot(props) +
    stat_lineribbon(aes(x = r, y = .value, col = .category), fill = "gray") +
    labs(x = "Distance from center", y = "Proportion"))
  ggsave(filename = str_glue("./hypothesis/figure_3/val/{name}_max_prob_distances.png"), plot = validation_plot, width = 8, height = 5)
  
  r_vals <- props %>%
    ungroup() %>%
    group_by(.category, r) %>%
    summarise(median_prob = median(.value)) 
  
  max_r <- map_df(c("rod", "transitioning", "spore"), function(label) {
    props <- filter(r_vals, .category == label)
    max_row <- which.max(props$median_prob)
    props[max_row,]
  })
  
  max_r %>%
    mutate(metadata = name) %>%
    write_csv(str_glue("./hypothesis/figure_3/maximum_radial_proportions/{name}_max_prob_distances.csv"))
  
  contrast <- pmap_dfr(max_r, function(...) {
    ttt <- tibble(...)
    td <- props %>%
      filter(r %in% unique(c(1, ttt$r, 60))) %>%
      compare_levels(variable = .value, by = r) %>%
      filter(.category == ttt$.category) %>%
      mutate(run_name = name)
    
    map_df(td$r %>% unique(), ~ hypothesis(filter(td, r == .x), c(".value > 0", ".value < 0"))$hypothesis %>% mutate(r_comparison = .x)) %>%
      bind_cols(ttt) %>%
      mutate(metadata = name) %>%
      write_csv(str_glue("./hypothesis/figure_3/figure_3_{name}_{ttt$.category}_hypothesis.csv"))
  }) %>%
    mutate(name = name)
  return(contrast)
}

all_contrasts <- map_df(list.files("./rds/model_rds/radius_label", pattern = ".*\\.rds", full.names = T),
                        ~ compare_radial_prop(.x, str_extract(.x, "(?<=/)[A-Za-z_0-9\\.]*(?=\\.rds$)"), "body_01")) %>%
  mutate(time_point = str_extract(name, "\\d\\d(?=h)"),
         labels_factor = factor(.category, levels = c("rod", "transitioning", "spore")))
```

Figure 3A -- Radial distribution of cell classifications
Merge radial proportions of all bodies (no body effect)
```{r}
all_models <- read_rds("./rds/model_rds/merged_radial_proportions_all_bodies.rds") %>%
  mutate(label = factor(.category, levels = c("rod", "transitioning", "spore")),
         label = factor(label, levels = c("rod", "transitioning", "spore")),
         time_point = str_replace(time_point, "h", " h"))

all_densities_data <- read_rds("./rds/model_rds/merged_density_models_radius_all_bodies.rds") %>%
  group_by(time_point, left_pt) %>%
  mutate(left_pt_90 = quantile(.value, 0.90), left_pt_10 = quantile(.value, 0.1)) %>%
  ungroup() %>%
  group_by(time_point)

all_densities <- all_densities_data %>%
  mutate(scaled_estimate = (.value - min(left_pt_10)) / (max(left_pt_90) - min(left_pt_10)),
         time_point = str_replace(time_point, "h", " h"))

(radial_proportion_plot <- ggplot(all_models) +
    stat_lineribbon(aes(x = r, y = .value, col = label, fill = label), .width = 0.90, alpha = 0.5, point_interval = median_hdci) +
    stat_lineribbon(data = all_densities, aes(x = left_pt, y = scaled_estimate, col = "density"),
                    fill = "darkgray", .width = 0.90, point_interval = median_hdci, alpha = 0.5) +
    labs(x = "Distance from center (μm)",
         y = "Proportion/Norm. Density",
         col = "", fill = "") +
    coord_cartesian(ylim = c(0, 1)) +
    facet_grid(. ~ time_point) +
    guides(fill = "none", color = guide_legend(override.aes = list(fill = NA))) + # Have to do some acrobatics for this plot: https://stackoverflow.com/a/21066173/11460621
    scale_color_manual(values = c("rod" = "#f8766d", "transitioning" = "#00ba38", "spore" = "#83b0fc", "density" = "black")) + 
    scale_fill_manual(values = c("rod" = "#f8766d", "transitioning" = "#00ba38", "spore" = "#83b0fc", "0.9" = "darkgrey")) +
    theme(legend.position = "top", axis.text = element_text(size=18), legend.text = element_text(size=18), axis.title = element_text(size=18), strip.text = element_text(size=18)))

ggsave("./figure_3/figure_3A_radial_proportions_model.png", bg = "white", plot = radial_proportion_plot, width = 12, height = 3.75)
```



Figure 4C + 4D Fluorescences
Fit model
```{r}
cell_data_expr <- read_rds("./rds/myxo_data.rds") %>%
  ungroup() %>%
  filter(body_id %in% c("body_19", "body_20", "body_21",
                        "body_22", "body_23", "body_24"),
         time_point != "24h") %>%
  group_by(body_id, time_point) %>%
  mutate(promoter = ifelse(body_id %in% c("body_19", "body_20", "body_21"), "dev", "fmgE")) %>%
  mutate(ch1_mean_log = MeanIntensity %>% log10,
         ch2_mean_log = channel_2_MeanIntensity %>% log10,
         time_point = ordered(time_point, c("27h", "30h", "36h", "48h"))) %>%
  ungroup() %>%
  filter((ch1_mean_log > log10(50) | ch2_mean_log > log10(100)))

mv_formula <- bf(mvbind(ch1_mean_log, ch2_mean_log) ~ promoter + labels_factor + mo(time_point):promoter:labels_factor +
                   (promoter + labels_factor + mo(time_point):promoter:labels_factor || body_id))

generic_prior <- "student_t(3, 0, 3)"

(mv_prior <- get_prior(mv_formula + set_rescor(FALSE), data = cell_data_expr))
mv_prior$prior[mv_prior$class == "sigma"] = "exponential(2)"
mv_prior$prior[mv_prior$class == "sd"] = "exponential(2)"
mv_prior$prior[mv_prior$prior == "" & mv_prior$class != "simo"] = generic_prior
mv_prior$prior[mv_prior$resp == "" ] = ""
mv_prior$prior[mv_prior$prior == generic_prior & mv_prior$coef == ""] = ""
(mv_prior)

mv_model <- brm(mv_formula + set_rescor(FALSE), data = cell_data_expr, chains = 8, cores = 8, inits = 0, prior = mv_prior,
                warmup = 250, iter = 750, control = list("adapt_delta" = 0.99, "metric" = "diag_e", "max_treedepth" = 20))

write_rds(mv_model, "./rds/model_rds/fluor_mv_model.rds", compress = "gz")
```

Figure 4C + 4D Fluorescences
Contrast, hypothesis, plot.
```{r}
mv_model <- read_rds("./rds/model_rds/fluor_mv_model.rds")

# mot = monotonically ordered time. Looking at the time "slope" from the model.
mot_vals <- spread_draws(mv_model, regex = T, `bsp_.*`) %>%
  mutate(dev_rod = `bsp_ch2meanlog_promoterdev:labels_factorrod:motime_point` - `bsp_ch1meanlog_promoterdev:labels_factorrod:motime_point`,
         fmgE_rod = `bsp_ch2meanlog_promoterfmgE:labels_factorrod:motime_point` - `bsp_ch1meanlog_promoterfmgE:labels_factorrod:motime_point`,
         dev_transitioning = `bsp_ch2meanlog_promoterdev:labels_factortransitioning:motime_point` - `bsp_ch1meanlog_promoterdev:labels_factortransitioning:motime_point`,
         fmgE_transitioning = `bsp_ch2meanlog_promoterfmgE:labels_factortransitioning:motime_point` -`bsp_ch1meanlog_promoterfmgE:labels_factortransitioning:motime_point`,
         dev_spore = `bsp_ch2meanlog_promoterdev:labels_factorspore:motime_point` - `bsp_ch1meanlog_promoterdev:labels_factorspore:motime_point`,
         fmgE_spore = `bsp_ch2meanlog_promoterfmgE:labels_factorspore:motime_point` - `bsp_ch1meanlog_promoterfmgE:labels_factorspore:motime_point`) %>%
  select(.chain, .iteration, .draw, dev_rod, fmgE_rod, dev_transitioning, fmgE_transitioning, dev_spore, fmgE_spore)

mot_vals_wide <- pivot_longer(mot_vals, dev_rod:fmgE_spore, "promoter_label")

mot_vals_hypothesis <- map_df(unique(mot_vals_wide$promoter_label), function(x) {
  tt <- filter(mot_vals_wide, promoter_label == x)
  th <- hypothesis(tt, hypothesis = "value > 0", alpha = 0.05)$hypothesis %>%
    mutate(promoter_label = x)
})s
write_csv(mot_vals_hypothesis, file = "./csv/hypothesis/figure_4C_mot_hypothesis.csv")

# Figure 4D
f4d_mot_comparison <- map_df(c("dev_transitioning", "dev_spore", "fmgE_transitioning", "fmgE_spore"), function(x) {
  ref <- ifelse(str_detect(x, "dev"), "dev_rod", "fmgE_rod")
  tt <- filter(mot_vals_wide, promoter_label %in% c(ref, x))
  comp <- compare_levels(tt, variable = value, by = promoter_label) %>%
    rename(comparison = promoter_label) %>%
    mutate(promoter_label = x)
})

f4d_mot_hypothesis <- map_df(unique(f4d_mot_comparison$comparison), function(x) {
  tt <- filter(f4d_mot_comparison, comparison == x)
  th <- hypothesis(tt, "value > 0", alpha = 0.05)$hypothesis %>%
    mutate(comparison = x)
})

write_csv(f4d_mot_hypothesis, file = "./csv/hypothesis/figure_4D_mot_hypothesis.csv")

(plotting_y_data <- cell_data_expr %>%
    select(labels:ch2_mean_log) %>%
    pivot_longer(cols = c(ch1_mean_log, ch2_mean_log), names_to = ".category") %>%
    mutate(.category = str_replace_all(.category, "_", ""),
           time_point = str_extract(label_time, "\\d\\dh")))

(f4c_data <- data_grid(cell_data_expr, time_point, promoter, labels_factor) %>%
    add_fitted_draws(mv_model, re_formula = NA) %>%
    compare_levels(variable = .value, by = .category))

(mv_contrasts <- f4c_data %>%
    ggplot() +
    stat_pointinterval(aes(x = time_point, y = .value, col = labels_factor), .width = 0.90, point_interval = median_hdci) +
    facet_grid(promoter ~ labels_factor) +
    labs(y = expression(log[10](relative~fluorescence~intensity)), x = "Hours PS", fill = "", col = "") +
    scale_x_discrete(labels = c("27", "30", "36", "48")) +
    theme(text = element_text(size = 18),
          axis.text = element_text(size = 18),
          strip.text = element_text(size = 18),
          strip.text.y = element_text(size = 18, face = "italic"),
          legend.text = element_text(size=18),
          legend.position = "top"))

ggsave(filename = "./figure_4/figure_4C_mv_fluor_model_contrast.png", mv_contrasts, width = 6, height = 6, bg = "white")

f4c_hypothesis <- map_df(cross2(unique(f4c_data$labels_factor), unique(f4c_data$promoter)), function(v) {
  td <- filter(f4c_data, labels_factor == v[[1]], promoter == v[[2]])
  tc <- td %>%
    compare_levels(variable = .value, by = time_point, comparison = "pairwise")

  th <- map_df(unique(tc$time_point), ~hypothesis(filter(tc, time_point == .x), hypothesis = ".value > 0")$hypothesis %>%
                 mutate(time_point = .x, labels_factor = v[[1]], promoter = v[[2]]))
  return(th)
})

write_csv(f4c_hypothesis, file = "./csv/hypothesis/figure_4C_hypothesis.csv")

f4c_hypothesis_between_labels <- map_df(cross2(unique(f4c_data$time_point), unique(f4c_data$promoter)), function(v) {
  td <- filter(f4c_data, time_point == v[[1]], promoter == v[[2]])
  
  tc <- td %>%
    compare_levels(variable = .value, by = labels_factor, comparison = "pairwise")

  th <- map_df(unique(tc$labels_factor), ~hypothesis(filter(tc, labels_factor == .x), hypothesis = ".value > 0")$hypothesis %>%
                 mutate(time_point = v[[1]], labels_factor = .x, promoter = v[[2]]))
  return(th)
})

write_csv(f4c_hypothesis_between_labels, file = "./csv/hypothesis/f4c_hypothesis_between_labels.csv")

(mv_model_fitted <- data_grid(cell_data_expr, time_point, promoter, labels_factor) %>%
    add_fitted_draws(mv_model, re_formula = NA) %>%
    compare_levels(variable = .value, by = .category, comparison = control))

mv_model_27h_rod <- mv_model_fitted %>% 
  filter(time_point == "27h", labels_factor == "rod")

mv_model_27h_rod_duplicated <- bind_rows(
  mv_model_27h_rod,
  mv_model_27h_rod %>% mutate(time_point = "30h"),
  mv_model_27h_rod %>% mutate(time_point = "36h"),
  mv_model_27h_rod %>% mutate(time_point = "48h")
)

f4d <- mv_model_fitted %>%
  filter(labels_factor != "rod") %>%
  bind_rows(mv_model_27h_rod_duplicated) %>%
  compare_levels(variable = .value, by = labels_factor, comparison = control) %>%
  mutate(labels_factor = str_replace(labels_factor, " - rod", ""),
         labels_factor = factor(labels_factor, levels = c("transitioning", "spore")))

(time_contrast_27hrod <- f4d %>%
    ggplot() +
    stat_pointinterval(aes(x = time_point, y = .value, col = labels_factor), .width = 0.90,
                       position = position_dodge(0.25), point_interval = median_hdci) +
    facet_grid(. ~ promoter) +
    geom_hline(yintercept = 0, linetype = 2) +
    labs(x = "Hours PS", y = expression(log[10](relative~fluorescence~intensity)), fill = "", col = "") +
    scale_x_discrete(labels = c("27", "30", "36", "48")) +
    theme(text = element_text(size = 18),
          axis.text = element_text(size = 18),
          strip.text.x = element_text(size = 18, face = "italic"),
          legend.text = element_text(size=18),
          legend.position = "top") +
    scale_color_manual(values = c("#00BA38", "#619CFF")))
ggsave(filename = "./figure_4/figure_4D_fluor_model_promoter_contrast_time_contrast_27h_rods.png", time_contrast_27hrod, width = 6, height = 6, bg = "white")

f4d_hypothesis <- map_df(cross3(unique(f4d$time_point), unique(f4d$labels_factor), unique(f4d$promoter)), function(v) {
  tt <- filter(f4d, time_point == v[[1]], labels_factor == v[[2]], promoter == v[[3]])
  th <- hypothesis(tt, hypothesis = ".value > 0")$hypothesis %>%
    mutate(time_point = v[[1]], labels_factor = v[[2]], promoter = v[[3]])
})

write_csv(f4d_hypothesis, path = "./csv/hypothesis/figure_4D_hypothesis.csv")

f4d_bewteen_hypothesis <- map_df(cross2(unique(f4d$time_point), unique(f4d$promoter)), function(v) {
  tt <- filter(f4d, time_point == v[[1]], promoter == v[[2]]) %>%
    compare_levels(variable = .value, by = labels_factor)
  th <- hypothesis(tt, hypothesis = ".value > 0")$hypothesis %>%
    mutate(time_point = v[[1]], promoter = v[[2]], comparison = tt$labels_factor[1])
  return(th)
})

write_csv(f4d_bewteen_hypothesis, path = "./csv/hypothesis/figure_4D_between_label_hypothesis.csv")
```


Figure 5A -- Radial distribution of fluorescence
pDev (FB19-FB21), pFmgE (FB22-FB24).
```{r}
cells_dist <- read_rds("./rds/myxo_data.rds") %>%
  filter(body_id %in% c("body_19", "body_20", "body_21", "body_22", "body_23", "body_24")) %>%
  mutate(promoter = ifelse(body_id %in% c("body_19", "body_20", "body_21"), "dev", "fmgE"),
         ch1_mean_log = MeanIntensity %>% log10(),
         ch2_mean_log = channel_2_MeanIntensity %>% log10(),
         relative_signal = (channel_2_MeanIntensity / MeanIntensity) %>% log10()) %>%
  filter((ch1_mean_log > log10(50) | ch2_mean_log > log10(100))) %>%
  mutate(promoter_by_time = str_c(promoter, time_point %>% as.character(), sep = "."),
         time_point_str = str_replace(time_point, "h", " h"),
         dummy = as.numeric(1.0))

gam_condition <- function(filtered_data, label_str) {
  print(label_str)
  
  formula <- bf(mvbind(ch1_mean_log, ch2_mean_log) ~ labels_factor + s(dist_from_center, by = labels_factor) +
                  s(dist_from_center, body_id, labels_factor, dummy, bs='re') + (labels_factor|body_id),
                sigma ~ labels_factor + s(dist_from_center, by = labels_factor) +
                  s(dist_from_center, body_id, labels_factor, dummy, bs='re') + (labels_factor|body_id)) + set_rescor(FALSE)
  
  (pr <- get_prior(formula, data = filtered_data))
  pr$prior[pr$group == "body_id"] = "exponential(0.5)"
  pr$prior[1] = "student_t(3, 0, 2.5)"
  
  fit <- brm(formula, data = filtered_data, cores = 4, chains = 4, warmup = 250,
             iter = 750, inits = 0, control = list("max_treedepth" = 15, "metric" = "diag_e", "adapt_delta" = 0.99))
  
  write_rds(fit, path = str_glue("./rds/model_rds/radius_intensity/{label_str}.rds"), compress = "gz")
}

future_walk2(list("27h", "27h",
          "30h", "30h",
          "36h", "36h",
          "48h", "48h"),
     list("dev", "fmgE",
          "dev", "fmgE",
          "dev", "fmgE",
          "dev", "fmgE"),
     ~ filter(cells_dist, time_point == .x, promoter == .y) %>% gam_condition(str_glue("gam_fluor_{.x}_{.y}")))
```

Figure 5A -- Radial distribution of fluorescence
All fruiting bodies (no body effect)
```{r}
merge_radial_intensity <- function(filepath, name, in_body_id) {
  print(str_glue("Loading {filepath}"))
  fit <- read_rds(filepath)
  
  # This is a weird workaround for a weird error introduced by the GAM
  body_group <- if (in_body_id %in% c("body_19", "body_20", "body_21")) {
    c("body_19", "body_20", "body_21")
  } else {
    c("body_22", "body_23", "body_24")
  }
  
  pred_no_body <- data_grid(fit$data, dist_from_center = seq(0, 60, 0.2), body_id = in_body_id %>% factor(levels = body_group), dummy = 0, labels_factor = c("rod", "transitioning", "spore")) %>%
    filter(body_id == in_body_id) %>%
    add_fitted_draws(fit, n = 1000) %>%
    mutate(name = name,
           promoter = str_extract(name, "(dev|fmgE)"),
           time_point = str_extract(name, "\\d{2}h"),
           labels_factor = factor(labels_factor, levels = c("rod", "transitioning", "spore")))
  
  return(pred_no_body)
}

body_lut <- c("dev" = "body_19", "fmgE" = "body_22")
all_models <- map_df(list.files("./rds/model_rds/radius_intensity", pattern = ".*fluor.*\\.rds$", full.names = T),
                     ~ merge_radial_intensity(.x, str_extract(.x, "(?<=/)[A-Za-z_0-9]*(?=\\.rds$)"),
                                             body_lut[str_extract(.x, "(dev|fmgE)")] %>% unlist))
write_rds(all_models, file = "./rds/model_rds/merged_intensity_models.rds", compress = "gz")
```


Figure 5A -- Radial distribution of fluorescence
All fruiting bodies (no body effect)
```{r}
all_models <- read_rds("./rds/model_rds/merged_intensity_models.rds") %>%
  mutate(labels_factor = factor(labels_factor, levels = c("rod", "transitioning", "spore")),
         time_point_str = str_replace(time_point, "h", " h"))

cell_data <- read_rds("./rds/myxo_data.rds") %>%
  filter(body_id %in% c("body_19", "body_20", "body_21", "body_22", "body_23", "body_24")) %>%
  mutate(dummy = 1.0,
         promoter = ifelse(body_id %in% c("body_19", "body_20", "body_21"), "dev", "fmgE"),
         ch1_log_mean = MeanIntensity %>% log10(),
         relative_signal = (channel_2_MeanIntensity / MeanIntensity) %>% log10())

all_models_contrast <- all_models %>%
  ungroup() %>%
  group_by(dist_from_center, labels_factor, time_point, promoter) %>%
  compare_levels(variable = .value, by = .category, comparison = list(c("ch2meanlog", "ch1meanlog"))) %>%
  mutate(time_point_str = str_replace(time_point, "h", " h"))

# Relative plot no rods
(radial_intensity_plot_norods <- all_models_contrast %>% 
    filter(labels_factor != "rod") %>%
    ggplot() +
    stat_lineribbon(aes(x = dist_from_center, y = .value, col = labels_factor, fill = labels_factor), alpha = 0.33, .width = 0.90, point_interval = median_hdci) +
    labs(x = "Distance from center (μm)",
         y = expression(log[10](relative~fluorescence~intensity)), col = "", fill = "") +
    facet_grid(promoter ~ time_point_str) +
    scale_color_manual(values = c("transitioning" = "#00ba38", "spore" = "#83b0fc")) + 
    scale_fill_manual(values = c("transitioning" = "#00ba38", "spore" = "#83b0fc")) +
    guides(fill = "none") +
    theme(text = element_text(size = 18),
          axis.text = element_text(size = 18),
          strip.text.x = element_text(size = 18),
          strip.text.y = element_text(size = 18, face = "italic"),
          legend.position = "top"))
ggsave(filename = "./figure_5/fig5_radial_intensity_relative.png", plot = radial_intensity_plot_norods, width = 12, height = 5.25, bg = "white")
```


Figure 5A -- Radial distribution of fluorescence
Contrasts all bodies (no body effect)
```{r}
fluorescent_gam_contrast <- function(filepath, name, body_id) {
  print(str_glue("Loading {filepath}"))
  fit <- read_rds(filepath)
  
  # This is a workaround for a weird error introduced by the GAM
  body_group <- if (fit$data$body_id[1] %in% c("body_19", "body_20", "body_21")) {
    c("body_19", "body_20", "body_21")
  } else {
    c("body_22", "body_23", "body_24")
  }
  
  # find point of maximal prob
  tt <- data_grid(fit$data, dist_from_center = seq(1, 60, 1), body_id = fit$data$body_id[1] %>% factor(levels = body_group), labels_factor = labels_factor, dummy = 0) %>%
    filter(body_id == fit$data$body_id[1]) %>%
    add_fitted_draws(fit, n = 1000, re_formula = NA)
  
  relative_fluor <- tt %>%
    ungroup() %>%
    group_by(dist_from_center, labels_factor, .category) %>%
    compare_levels(variable = .value, by = .category, comparison = list(c("ch2meanlog", "ch1meanlog")))
  
  r_vals <- relative_fluor %>%
    ungroup() %>%
    group_by(.category, labels_factor, dist_from_center) %>%
    summarise(median_val = median(.value)) 
  
  max_r <- map_df(c("rod", "transitioning", "spore"), function(label) {
    tt <- filter(r_vals, labels_factor == label)
    max_row <- which.max(tt$median_val)
    tt[max_row,]
  })
  
  max_r %>% mutate(metadata = name) %>% write_csv(str_glue("./csv/fluorescence_hypotheses/max_distances/{name}_max_fluor_distances.csv"))
  
  td <- relative_fluor %>%
    filter(dist_from_center == 1 || dist_from_center == 30 || dist_from_center == 60) %>%
    compare_levels(variable = .value, by = dist_from_center) %>%
    mutate(run_name = name)
  
  map_df(cross2(td$dist_from_center %>% unique(), c("rod", "transitioning", "spore")), ~ hypothesis(filter(td, dist_from_center == .x[[1]], labels_factor == .x[[2]]),
                                                        c(".value > 0", ".value < 0"))$hypothesis %>% mutate(r_comparison = .x[[1]], labels_factor = .x[[2]])) %>%
    mutate(metadata = name) %>%
    write_csv(str_glue("./csv/fluorescence_hypotheses/figure_5_{name}_hypothesis.csv"))
}

all_contrasts <- walk(list.files("./rds/model_rds/radius_intensity", pattern = ".*\\.rds", full.names = T),
                        ~ fluorescent_gam_contrast(.x, str_extract(.x, "(?<=/)[A-Za-z_0-9\\.]*(?=\\.rds$)"), "body_01"))
```
